---
title: "Initial EDA Basic Outline"
author: "Arnav Bedi"
output: html_notebook
---

This is an outline for your Initial EDA. You can follow this outline, or you can modify it as you see fit. Do what works best for your project. 

## Research Question 

My topic for this final project is going to be Analyzing the Effect of Social Media on Mental Health

I want to explore the relationship between social media usage and mental health outcomes. I want to research this and make a project on this topic as it would be beneficial in the identification of the risk factors, promotion of positive usage of social media as well as these findings can be shared with mental health professionals, educators, and social media companies to help them understand how to better support users’ mental health. By analyzing data from social media platforms and mental health surveys, I can also identify correlations between screen time, content type, and mental health indicators such as anxiety, depression, and well-being. 



## Data Processing 
Installing neccesary libraries 

```{r}
# load necessary packages
library(tidyverse)
library(mosaic)
library(esquisse)

devtools::install_github("mdbeckman/dcData")
library(dcData)
```

#### Data Intake 

Social Media data set:
```{r}

social_media <- read.csv("C:/Users/Arnav/Downloads/val.csv")
head(social_media)

colnames(social_media)
```
Mental Health data set:
```{r}

mental_health <- read.csv("C:/Users/Arnav/Downloads/smmh.csv")
head(mental_health)

colnames(mental_health)

```
#### Data Wrangling 

Include R code and written explanation for wangling your data (you can make multiple wrangled data sets). 

```{r}
# Inspect the structure of the datasets
str(social_media)
str(mental_health)

# Checking for missing values
summary(social_media)
summary(mental_health)

# Ensuring that Daily_Usage_Time..minutes. column is numeric
social_media$Daily_Usage_Time..minutes. <- as.numeric(as.character(social_media$Daily_Usage_Time..minutes.))

# Removing rows where platform is "Female" or "55"
social_media <- social_media %>%
  filter(Platform != "Female", Platform != "55")

# Removing rows with non-numeric values in daily_usage_time_minutes
social_media <- social_media %>%
  filter(!is.na(as.numeric(Daily_Usage_Time..minutes.)), !is.na(Platform)) %>%
  mutate(Daily_Usage_Time..minutes. = as.numeric(Daily_Usage_Time..minutes.))

# Checking for conversion issues
summary(social_media$Daily_Usage_Time..minutes.)

# Aggregating social media usage per day
social_media_summary <- social_media %>%
  group_by(User_ID) %>%
  summarise(total_hours_spent = sum(Daily_Usage_Time..minutes., na.rm = TRUE) / 60) # Converting minutes to hours

# Aggregating average daily usage time per platform
platform_summary <- social_media %>%
  group_by(Platform) %>%
  summarise(avg_daily_usage_time = mean(Daily_Usage_Time..minutes., na.rm = TRUE) / 60)

# Aggregating mental health scores per day
mental_health_summary <- mental_health %>%
  group_by(Timestamp) %>%
  summarise(avg_depression_score = mean(X18..How.often.do.you.feel.depressed.or.down., na.rm = TRUE),
            avg_anxiety_score = mean(X13..On.a.scale.of.1.to.5..how.much.are.you.bothered.by.worries., na.rm = TRUE),
            daily_social_media_time = X8..What.is.the.average.time.you.spend.on.social.media.every.day.,)

# Aggregating average daily usage time per dominant emotion
emotion_summary <- social_media %>%
  group_by(Dominant_Emotion) %>%
  summarise(avg_daily_usage_time = mean(Daily_Usage_Time..minutes., na.rm = TRUE) / 60)


```


## Data Visualization

#### Plot 1

##Plot for Average Daily Usage Time per Platform

```{r}

ggplot(platform_summary, aes(x = Platform, y = avg_daily_usage_time, fill = Platform)) +
  geom_bar(stat = 'identity') +
  labs(title = "Average Daily Usage Time per Platform (hours)",
       x = "Platform",
       y = "Average Daily Usage Time") +
  theme_minimal()
```
This bar chart shows the average daily usage time on different social media platforms. Each bar represents a platform, and the height of the bar indicates the average daily usage time in minutes.

##Plot average daily usage time per dominant emotion:

```{r}

ggplot(emotion_summary, aes(x = Dominant_Emotion, y = avg_daily_usage_time, fill = Dominant_Emotion)) +
  geom_bar(stat = 'identity') +
  labs(title = "Average Daily Usage Time per Dominant Emotion (hours) ",
       x = "Dominant Emotion",
       y = "Average Daily Usage Time") +
  theme_minimal()

```

#### Plot 2

## Plot for Relationship between Social Media Usage and Mental Health Indicators


```{r}

ggplot(mental_health_summary, aes(x = as.factor(daily_social_media_time), y = avg_depression_score)) +
  geom_bar(stat = 'identity', fill = "blue", alpha = 0.7) +
  labs(
    title = "Average Depression Score by Daily Social Media Usage",
    x = "Daily Social Media Usage (hours)",
    y = "Average Depression Score"
  ) +
  theme_minimal() +
  scale_x_discrete(breaks = as.factor(mental_health_summary$daily_social_media_time)) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    plot.title = element_text(size = 16, hjust = 0.5)
  )

```
The bar graph illustrates the relationship between daily social media usage and the average depression score reported by participants. The x-axis represents the daily time spent on social media, measured in minutes, while the y-axis shows the average score for feeling depressed.

## Conclusion

#Initial Intuition for the plots in Plot 1:

-Platform Comparison: Some platforms might show higher average depression scores than others. For example, users of certain platforms might report higher feelings of depression due to different usage patterns or social dynamics.

-Social Media Influence: If a particular platform is associated with higher average depression scores, this could suggest that the platform’s characteristics (ex., user interactions, content type) might contribute to mental health outcomes.


#Initial Intuition for the plot in Plot 2:

-Usage Impact: The graph might show whether higher daily social media usage is associated with higher average depression scores. If the line trends upwards, it would suggest that increased social media use correlates with more frequent feelings of depression.

-Pattern Identification: Observing any distinct patterns or thresholds (e.g., after a certain amount of daily usage, depression scores significantly increase) could help in understanding how social media usage impacts mental health.
