---
title: "Initial EDA Basic Outline"
author: "Arnav Bedi"
output: html_notebook
---

This is an outline for your Initial EDA. You can follow this outline, or you can modify it as you see fit. Do what works best for your project. 

## Research Question 

My topic for this final project is going to be Analyzing the Effect of Social Media on Mental Health

I want to explore the relationship between social media usage and mental health outcomes. I want to research this and make a project on this topic as it would be beneficial in the identification of the risk factors, promotion of positive usage of social media as well as these findings can be shared with mental health professionals, educators, and social media companies to help them understand how to better support users’ mental health. By analyzing data from social media platforms and mental health surveys, I can also identify correlations between screen time, content type, and mental health indicators such as anxiety, depression, and well-being. 



## Data Processing 
Installing neccesary libraries 

```{r}
library(tidyverse)
library(mosaic)
install.packages("ggplot2")


```

#### Data Intake 

Social Media data set:
```{r}

social_media <- read.csv("C:/Users/Arnav/Downloads/val.csv")
head(social_media)


```
Mental Health data set:
```{r}

mental_health <- read.csv("C:/Users/Arnav/Downloads/smmh.csv")
head(mental_health)

```
#### Data Wrangling 

Include R code and written explanation for wangling your data (you can make multiple wrangled data sets). 

```{r}
# Inspect the structure of the datasets
str(social_media)
str(mental_health)

# Check for missing values
summary(social_media)
summary(mental_health)

social_media_summary <- social_media %>%
  group_by(User_ID) %>%
  summarise(avg_social_media_usage = mean(hours_spent, na.rm = TRUE))

```


## Data Visualization

#### Plot 1

Include R code to make one data visualization (I would encourage you to use 3 or more variables). Include a brief written description of what your graph shows. Is this graph helpful in answering your research question? If it is helpful, will you need to add/modify anything in this graph in the final report? 

#### Plot 2

Include R code to make one data visualization (I would encourage you to use 3 or more variables). Include a brief written description of what your graph shows. Is this graph helpful in answering your research question? If it is helpful, will you need to add/modify anything in this graph in the final report? 


## Conclusion

Include your initial intuition as to what the answer to your research question is using the data sets and visualizations you created. 
